---
title: "Reproducible Research (Course Project 1)"
author: "Christopher Johnson"
date: "November 18, 2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Set working driectory  

```{r}
setwd("\\\\cscphome/HOME1/CJOHNSO2/Redir-cjohnso2/Documents/RR Proj 1")
```
Load in library 
```{r}
library(reshape2)
library(knitr)
```

Load in data  

```{r}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
        temp <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
        unlink(temp)
}
activity <- read.csv("activity.csv")
```
View Data  
```{r}
head(activity)
```
###What is mean total number of steps taken per day?
Sum steps by day, create Histogram.  
add a curve line for steps taken  
```{r First_plot, fig.path = "figures/"}
meltedActivity <- melt(activity, id=c("date"), na.rm=TRUE, measure.vars="steps")
steps_by_day <- dcast(meltedActivity, date ~ variable, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", border = "blue", xlab="Number of Steps", prob= TRUE)
lines(density(steps_by_day$steps),col= "blue", lwd=3)
```

Calculate and report the mean and median of the total number of steps taken per day  
```{r}
actMean <- format(round(mean(steps_by_day$steps), 2), nsmall = 2)
actMedian <- median(steps_by_day$steps)
actMean
actMedian
```

##What is the average daily activity pattern?
*Calculate average steps for each interval for all days. 
*Plot the Average Number Steps per Day by Interval. 

```{r 2nd Plot, fig.path= "figures/"}
steps_by_interval <- aggregate(steps ~ interval, activity, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

Now which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
Max_steps <- steps_by_interval[which.max(steps_by_interval$steps),1]
Max_steps
```

###Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)  

```{r}
na_values <- activity$steps
count_of_NA <- length(which(is.na(na_values)))
count_of_NA
```
Now attempt to add data to where the NA's value are and test if worked  
```{r}
activity2<- activity
nas<- is.na(activity2$steps)
avg_interval<- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify = TRUE)
activity2$steps[nas] <- avg_interval[as.character(activity2$interval[nas])]
names(activity2)
```
check to see if missing cvalues appear 
```{r}
sum(is.na(activity2))
```

Create Histogram of new Data with out NA values
Create Histogram to show difference. 
```{r 3rd Plot,fig.path= "figures/"}
hist(activity2$steps, main = paste("Total Steps Each Day"), col="green", border = "blue", xlab="Number of Steps")

hist(activity$steps, main = paste("Total Steps Each Day"), col="red", border = "blue", xlab="Number of Steps", add=T)
legend("topright", c("Non NA's", "NA's"), col=c("green", "red"), lwd=10)
```


Calculate and report the new mean and median of the total number of steps taken per day. 
```{r}
meltedActivity2 <- melt(activity, id=c("date"), na.rm=TRUE, measure.vars="steps")
steps_by_day2 <- dcast(meltedActivity2, date ~ variable, sum)
newmean <- format(round(mean(steps_by_day2$steps), 2), nsmall = 2)
newmedian <- median(steps_by_day2$steps)
newmean
newmedian
```
The impact of imputing missing data on the estimates of the total daily number of steps is the smae since na's were removed from the orignal calculation. 

###Are there differences in activity patterns between weekdays and weekends?
Create a plot to compare weeddays and weekends, to see if there isa difference of activity.

```{r Final Plot,fig.path= "figures/"}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
activity2$dow <- as.factor(ifelse(is.element(weekdays(as.Date(activity2$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, activity2, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")


```

